Neighbor-Based Decentralized Training Strategies for Multi-Agent Reinforcement Learning

### 2024-10-12 paper-2024-sac-marl-neighbouring-strategies.tex
paper restructure

In \Cref{alg:neighbor_dqn} we provide a detailed description of the integration of the neighboring-based strategies with deep Q-learning.
\begin{algorithm}
  \caption{Neighboring-Based Deep Q-Learning}
  \label{alg:neighbor_dqn}
  \ForEach{agent $i \in N$}{
      Initialize Q-network $Q^i(o^i, a^i; \theta^i)$ and target network $Q^{i-}(o^i, a^i; \theta^{i-})$\\
      Initialize replay buffer $\mathcal{D}^i$
  }
  \For{episode = 1 to M}{
      \ForEach{agent $i \in N$}{
          Initialize initial observation $o_0^i$
      }
      \For{t = 0 to T}{
          \ForEach{agent $i \in N$}{
              Select action $a_t^i$ based on $\epsilon$-greedy policy using $Q^i(o_t^i, a^i; \theta^i)$\\
              Execute action $a_t^i$ and observe reward $r_{t+1}^i$ and next observation $o_{t+1}^i$\\
              Store $(o_t^i, a_t^i, r_{t+1}^i, o_{t+1}^i)$ in $\mathcal{D}^i$\\
              Sample a mini-batch from $\mathcal{D}^i$\\
              Calculate target values using chosen neighboring strategy (see below)\\
              Update $\theta^i$ by minimizing the TD error\\
              Every C steps, update target network $\theta^{i-} \leftarrow \theta^i$\\
              \textbf{
              Every S steps, perform neighbor interaction (e.g., averaging, consensus, experience sharing)}
          }
      }
  }
\end{algorithm}
\section{Experimental setup}\label{sec:experiments}

To validate our approach, we developed a custom learning scenario using three state-of-the-art frameworks 
  in RL and DL, namely: \emph{Ray}~\cite{DBLP:conf/osdi/MoritzNWTLLEYPJ18} 
  (in particular, its module \emph{RLlib}~\cite{DBLP:conf/icml/LiangLNMFGGJS18}),
  \emph{Gymnasium}~\cite{DBLP:journals/corr/abs-2407-17032}
  and \emph{PyTorch}~\cite{DBLP:conf/nips/PaszkeGMLBCKLGA19}.

We ran a total of $35$ training experiments, 
namely: $5$ different methods %% TODO check how many ex are discussed
  (centralized, distributed and the three neighbor strategies discussed).
Each method was evaluated across 10 distinct random seeds to enhance the statistical robustness of the observed results.
%
Each experiment consisted of $50$ training episodes.

Furthermore, we ran $15$ evaluation experiments, namely: $5$ different methods and for each method 
  we varied the number of agents ($8$, $16$ and $32$) to assess the metrics under various conditions.
  In this case, for each experiment we ran $100$ evaluation episodes.

All the experiments are publicly available and reproducible\footnote{\url{https://anonymous.4open.science/r/experiments-2024-SAC-marl-neighbouring-strategies-4B8F}}.

\subsection{Simulated environment}
The proposed multi-agent environment, which we design to evaluate the efficacy of the neighboring-based strategies,
 consists of a two-dimensional Euclidean space 
 populated by a variable number of \emph{targets} and \emph{agents}, 
 whose initial positions are randomly distributed across the space.
%
The environment is considered solved when all targets are collected, terminating the episode. 
To prevent infinitely long episodes, a maximum time limit of T = 300 steps is imposed. 
If all items are collected before this limit, the episode terminates early.
The agents' objective  is to collaboratively collect all the targets in the environment, minimizing the total retrivial time.
%
The observation space of each agent consist of a tuple of three elements: %% TODO CHECK AGAIN
\begin{enumerate*}[label=(\roman*)]
  \item number of the closest agents that the agent is able to see (thus, forming a neighborhood);
  \item number of the closest items that the agent is able to see;
  \item memory of past states, thus encoding also temporal information, helping each agent to understand 
    the evolution of the environment.
\end{enumerate*} 
%
The action space for each agent is discrete, 
consisting of eight possible movements corresponding 
to the cardinal and diagonal directions on a grid.
%

@@

@@

\section{Results and Evaluation}\label{sec:results}

>>>>

In \Cref{alg:neighbor_dqn} we provide a detailed description of the integration of the neighboring-based strategies with deep Q-learning.
\begin{algorithm}
  \caption{Neighboring-Based Deep Q-Learning}
  \label{alg:neighbor_dqn}
  \ForEach{agent $i \in N$}{
      Initialize Q-network $Q^i(o^i, a^i; \theta^i)$ and target network $Q^{i-}(o^i, a^i; \theta^{i-})$\\
      Initialize replay buffer $\mathcal{D}^i$
  }
  \For{episode = 1 to M}{
      \ForEach{agent $i \in N$}{
          Initialize initial observation $o_0^i$
      }
      \For{t = 0 to T}{
          \ForEach{agent $i \in N$}{
              Select action $a_t^i$ based on $\epsilon$-greedy policy using $Q^i(o_t^i, a^i; \theta^i)$\\
              Execute action $a_t^i$ and observe reward $r_{t+1}^i$ and next observation $o_{t+1}^i$\\
              Store $(o_t^i, a_t^i, r_{t+1}^i, o_{t+1}^i)$ in $\mathcal{D}^i$\\
              Sample a mini-batch from $\mathcal{D}^i$\\
              Calculate target values using chosen neighboring strategy (see below)\\
              Update $\theta^i$ by minimizing the TD error\\
              Every C steps, update target network $\theta^{i-} \leftarrow \theta^i$\\
              \textbf{
              Every S steps, perform neighbor interaction (e.g., averaging, consensus, experience sharing)}
          }
      }
  }
\end{algorithm}
\section{Evaluation}\label{sec:experiments}
\subsection{Simulated environment}
The proposed multi-agent environment, which we design to evaluate the efficacy of the neighboring-based strategies,
 consists of a two-dimensional Euclidean space 
 populated by a variable number of \emph{targets} and \emph{agents}, 
 whose initial positions are randomly distributed across the space.
%
The environment is considered solved when all targets are collected, terminating the episode. 
To prevent infinitely long episodes, a maximum time limit of T = 300 steps is imposed. 
If all items are collected before this limit, the episode terminates early.
The agents' objective  is to collaboratively collect all the targets in the environment, minimizing the total retrivial time.
%
The observation space of each agent consist of a tuple of three elements: %% TODO CHECK AGAIN
\begin{enumerate*}[label=(\roman*)]
  \item number of the closest agents that the agent is able to see (thus, forming a neighborhood);
  \item number of the closest items that the agent is able to see;
  \item memory of past states, thus encoding also temporal information, helping each agent to understand 
    the evolution of the environment.
\end{enumerate*} 
%
The action space for each agent is discrete, 
consisting of eight possible movements corresponding 
to the cardinal and diagonal directions on a grid.
%

@@

\subsection{Experimental setup}
To validate our approach, we developed a custom learning scenario using three state-of-the-art frameworks 
  in RL and DL, namely: \emph{Ray}~\cite{DBLP:conf/osdi/MoritzNWTLLEYPJ18} 
  (in particular, its module \emph{RLlib}~\cite{DBLP:conf/icml/LiangLNMFGGJS18}),
  \emph{Gymnasium}~\cite{DBLP:journals/corr/abs-2407-17032}
  and \emph{PyTorch}~\cite{DBLP:conf/nips/PaszkeGMLBCKLGA19}.

We ran a total of $35$ training experiments, 
namely: $5$ different methods %% TODO check how many ex are discussed
  (centralized, distributed and the three neighbor strategies discussed).
Each method was evaluated across 10 distinct random seeds to enhance the statistical robustness of the observed results.
%
Each experiment consisted of $50$ training episodes. 
%
Furthermore, we ran $15$ evaluation experiments, namely: $5$ different methods and for each method 
  we varied the number of agents ($8$, $16$ and $32$) to assess the metrics under various conditions.
  In this case, for each experiment we ran $100$ evaluation episodes.
All the experiments are publicly available and reproducible\footnote{\url{https://anonymous.4open.science/r/experiments-2024-SAC-marl-neighbouring-strategies-4B8F}}.

@@

\subsection{Results}\label{sec:results}
